# Working on activity data

## Loading the File and reading the summary
```{r}
data = read.csv("C:\\Users\\nilak\\Coursera_DSTB\\activity.csv")
summary(data)
```

## Histogram of total steps taken each day

```{r}
totalsteps = aggregate(steps~date,data,sum)
names(totalsteps)[2] = 'totalsteps'
hist(totalsteps$totalsteps,xlab = 'Total Steps in a day',ylab = 'Frequency of total steps',main = 'Histogram of total steps', col = 'blue')
```

## Mean and Median of the data Steps

```{r}
print(c(mean(totalsteps$totalsteps),median(totalsteps$totalsteps)))
```

## Histogram of steps in different intervals

```{r}
intervalsteps = aggregate(steps~interval,data,mean)
names(intervalsteps)[2] = 'Avg steps'
hist(intervalsteps$`Avg steps`,xlab = "Average steps in the 5 min interval",ylab = 'Frequency of the avgsteps over all days',main = 'Histogram of the average steps in the 5 min interval',col = 'green')
```


## Time Series Plot of the steps taken in the five minute interval

```{r}
plot(intervalsteps$interval, intervalsteps$`Avg steps`, type="l", col="green" , lwd=2)
```

## Finding the interval that contains on average max steps

```{r}
intervalsteps[which.max(intervalsteps$`Avg steps`),1]
```

## Number of missing steps

```{r}
sum(is.na(data$steps))
```

## Filling missing values with mean of interval steps

```{r}
datafill = merge(x = data, y = intervalsteps, by = "interval", all.x = TRUE)
datafill$steps[is.na(datafill$steps)] = datafill$`Avg steps`[is.na(datafill$steps)]
sum(is.na(datafill))

```

## Histogram of total steps taken each day

```{r}
totalstepsfill = aggregate(steps~date,datafill,sum)
names(totalstepsfill)[2] = 'totalsteps'
hist(totalstepsfill$totalsteps,xlab = 'Total Steps in a day',ylab = 'Frequency of total steps',main = 'Histogram of total steps', col = 'red')

```

## Median and mean of the total steps taken each day after imputing missing values

```{r}
mean(totalstepsfill$totalsteps)
median(totalstepsfill$totalsteps)
```

## Creating the weekday variable

```{r}
datafill$date = as.Date(datafill$date)
datafill$week = weekdays(datafill$date)
datafill$weekfactor = ifelse(datafill$week == "Monday"|datafill$week == "Tuesday"|datafill$week == "Wednesday"|datafill$week == "Thursday"|datafill$week == "Friday" ,1,0)
```

## Time Series Plot of the average steps taken in the 5 minute intervals on weekdays and weekends 

```{r}
plot(intervalsteps$interval, intervalsteps$`Avg steps`, type="l", col="green" , lwd=2)
intervalsteps = aggregate(steps~interval+weekfactor,datafill,mean)
names(intervalsteps)[3] = 'Average'
names(intervalsteps)[2] = 'weekfactor'
library(lattice)
xyplot(Average~interval|factor(weekfactor),type='l', layout=c(1,2), xlab='Interval',ylab='Number of Steps',data = intervalsteps,col = "brown")

```




